---
title: "Conditional Variational Autoencoder"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### 1. Import libraries

```{r message=FALSE, warning=FALSE}
# With TF-2, you can still run this code due to the following line:
if (tensorflow::tf$executing_eagerly())
  tensorflow::tf$compat$v1$disable_eager_execution()

library(keras)
K <- keras::backend()

#library(keras)
library(tidyverse)
library(imager)
library(recolorize)
library(OpenImageR)
library(readxl)
```

##### 2. Import dataset

```{r}
files_train <- list.files("Dataset/Images/Train_filtered",  full.names = TRUE, pattern = ".jpg", all.files = TRUE)
#files_train_labels <- list.files("Dataset/Features/Features_Class_Train/",  full.names = TRUE, pattern = ".txt", all.files = TRUE)
labels_train <- read_excel('Dataset/Features/Features_Dummy_Train/Labels_Dummy.xlsx')

files_test <- list.files("Dataset/Images/Test_filtered",  full.names = TRUE, pattern = ".jpg", all.files = TRUE)
#files_test_labels <- list.files("Dataset/Features/Features_Class_Test/",  full.names = TRUE, pattern = ".txt", all.files = TRUE)
labels_test <- read_excel('Dataset/Features/Features_Dummy_Test/Labels_Dummy_test.xlsx')

```

##### 3. Preprocessing dataset

```{r}

```


This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
